Eddy covariance towers as sentinels of abnormal radioactive material releases

### 2024-11-19 manuscript/manuscript.tex
text edits

Our utilization of flux tower records for event discrimination stands in contrast to the typical use of data from the towers for direct carbon balance accounting. Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange \citep{baldocchiInterannualVariabilityNet2018}, our use treats each variable as potentially including an indirect signal (i.e. an anomaly from normal vegetation behavior) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc). In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way \citep{nobelPhysicochemicalEnvironmentalPlant2020}.

@@

We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other eddy flux towers. While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, prior studies support the idea that exposure may cause declines in photosynthetic $CO_2$ uptake. For example, \citet{gudkovEffectIonizingRadiation2019} detail how a wide range of exposure doses and durations can inhibit CO2 uptake as a result of enzymatic effects on the Calvin Cycle. For the present study, we lack leaf-level observations to directly establish inhibition and instead express it as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g. air temperature, humidity, precipitation, etc.) before, during, and after exposure.

@@

\section{Discussion}
Using flux tower records collected closest to and in the prevailing wind direction to the IRE and the Fukushima release locations, our event detection algorithm was able to identify both known abnormal releases and plausible previously unidentified abnormal events. There was broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases, and our approach was supported by the loss of signal at flux towers outside of the prevailing wind direction. This suggests that, at the very least, given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event. Furthermore, our approach has promise for situations when there is not a prior known abnormal release. In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study. The delay between data collection and availability to the general community via data repositories (e.g., Fluxnet) is highly variable among sites and investigators. In the best case scenario, for actively monitored sites with responsive investigators, this can be as little as 6 months. As a result, operationalization of our approach might target post-hoc attribution rather than real-time monitoring applications.

More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep[e.g., ][]{fratiniEddyCovarianceFlux2018}. We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records. This strategy of using only limited sections of the data is particularly important because it can avoid the situation where a well-defined linear relationship becomes non-linear and obscured at longer timescales. Despite the apparent success of our approach, it is not a perfect solution because some non-linearity remains. Beyond the noise inherent to most environmental data, this is likely one of the principal causes of the relatively modest $R^2$ values we observed in Table 1. These are not unusally low however as $R^2$ values between 0.35 and 0.7 are typical for environmental data \citep{murrayMethodsQuantifyVariable2009}. A fruitful area of future research would be exploring physical justifications for linearizing these relationships, which would likely lead to more precise signal recovery and event detection. Our approach was further aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers. This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal. As a result, the ability of our approach to identify a suitable release signature does depend on the quality and completeness of the underlying data.

>>>>

Our utilization of flux tower records for event discrimination stands in contrast to the typical use of data from the towers for direct carbon balance accounting. Whereas carbon balance accounting activities deal directly in the units of the measured data for computing quantities like net ecosystem exchange \citep{baldocchiInterannualVariabilityNet2018}, our use treats each variable as potentially including an indirect signal (i.e., an anomaly from normal vegetation behavior or expected carbon or water vapor fluxes) of radioactive material deposition irrespective of its physical properties (e.g., mass, volume, etc). In this way, we are not directly measuring radioactive deposition but rather leveraging the fact that existing ecosystem monitoring records reflect vegetation photosynthesis rates and general stress condition of site vegetation in a known way \citep{nobelPhysicochemicalEnvironmentalPlant2020}.

@@

We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other eddy flux towers. While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, prior studies support the idea that exposure may cause declines in photosynthetic $CO_2$ uptake. For example, \citet{gudkovEffectIonizingRadiation2019} detail how a wide range of exposure doses and durations can inhibit CO2 uptake as a result of enzymatic effects on the Calvin Cycle. For the present study, we lack leaf-level observations to directly establish inhibition and instead express it indirectly as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g. air temperature, humidity, precipitation, etc.) before, during, and after exposure. A consequence of this indirect approach is that we cannot directly separate material deposition responses coming from vegetation apart from possible deposition interference with flux tower sensors.

@@

\section{Discussion}
Using flux tower records collected closest to and in the prevailing wind direction to the IRE and the Fukushima release locations, our event detection algorithm was able to identify both known abnormal releases and plausible previously unidentified abnormal events. There was broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases, and our approach was supported by the loss of signal at flux towers outside of the prevailing wind direction. This suggests that, at the very least, given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event. Furthermore, our approach has promise for situations when there is not a prior known abnormal release. In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study. The delay between data collection and availability to the general community via data repositories (e.g., Fluxnet) is highly variable among sites and investigators. In the best case scenario, for actively monitored sites with responsive investigators, this can be as little as 6 months. As a result, operationalization of our approach might target post-hoc attribution rather than real-time monitoring applications unless release protocols of flux data are increased in frequency.

More generally, we show that flux tower records have value for this type of data mining despite the apparent noisiness of the data \citep[e.g., ][]{fratiniEddyCovarianceFlux2018}. We attribute the sensitivity of our approach to the fact that we fit successive interaction models to limited sections of the data record, and we leveraged bivariate relationships to constrain noisiness in individual data records. This strategy of using only limited sections of the data is particularly important because it can avoid the situation where a well-defined linear relationship becomes non-linear and obscured at longer timescales. Despite the apparent success of our approach, it is not a perfect solution because some non-linearity remains. Beyond the noise inherent to most environmental data, this is likely one of the principal causes of the relatively modest $R^2$ values we observed in Table 1. These are not unusally low however as $R^2$ values between 0.35 and 0.7 are typical for environmental data \citep{murrayMethodsQuantifyVariable2009}. A fruitful area of future research would be exploring physical justifications for including non-linear relationships, which would likely lead to more precise signal recovery and event detection. Our approach was further aided by the fact that this type of scale-free data mining is largely unaffected by systematic errors due to uncertainty in calibration standards and/or low resolution in specific sensor packages deployed on the towers. This is because although such systematic errors may bias the magnitude of overall flux values, which would be a problem for typical uses of flux towers that deal directly in the units of the measured data, they do not reduce the overall confidence in individual values \citep{langfordEddycovarianceDataLow2015}, which for our purposes might represent an abnormal release signal. As a result, the ability of our approach to identify a suitable release signature does depend on the quality and completeness of the underlying data.
